---
title: 'Homework 1: PSTAT 131'
author: "Sabrina Lem"
date: "3/30/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
***ML Main Ideas***
**Question 1:**
\
Supervised: Data used in supervised learning is labeled. As a data scientist,
we must use this labeled data where we know both the input and output
of the desired model to train the model to best map the given information.
\
\
Unsupervised: Unsupervised learning uses unlabeled data. Unsupervised
learning investigates patterns based on the given input data.
\
\
The main differences between supervised and unsupervised learning is that
in supervised learning we know the input and output of labeled data. 
While, in unsupervised learning you only know the input of unlabeled data.
Supervised learning branches into classification and regression problems,
and unsupervised learning can be separated into clustering and
association problems. 
\
\
**Question 2:**
\
In supervised learning, we can categorize the problems into regression and
categorical problems based on the nature of the response variable of 
the model.
\
Regression is used when the response variable is quantitative.
Classification is used when the response variable is qualitative.
\
\
**Question 3:**
\
Metrics for regression ML problems:
1. Mean Squared Error (MSE)
2. Mean Absolute Error
Metrics for classification ML problems
1.Confusion Matrix
2.Precision
\
\
**Question 4:**
\
Descriptive Model:A model that best visually reports the trend of
the data set.
Inferential Model:A model used to "asses the quality of our predictions and 
(or) estimation" (slide 30, lecture_day1). It is a model that "aims to test
theories" and "state the relationship between outcome & predictors" (slide 7
lecture_day2)
Prediction model: a model that uses past data on predictors to "accurately 
predict future response" (slide 30, day_1_131_231). It "aims to predict Y with
minimum reducible error" (slide 7, lecture_day2).
\
\
**Question 5:**
\
Mechanistic means that we "assume a parametric form of the function, $f$.
Empirically-driven means that we do not assume anything about $f$.
Mechanistic model can be made more flexible by adding parameters, however 
empirically-driven models are flexibly by default due to the fact that
the actual data points determine the function f (in an empirically driven 
model). That said, Empirically-driven models tend to require more
observations. Both models are at risk of over fitting. Both models are used
to made predictions; they are predictive models.
\
\
Empirically-driven models are easier to understand. This is because
it makes sense to form a model based on the actual observations, rather
than use theory to form a model (in the case of mechanistic models). 
\
\
Bias-variance trade off occurs when we want to decrease the variance of
the model or decrease the bias of a model. Whenever we  attempt to lower one
of these characteristics, the other will increase. In general, we see that
a simple model will have high bias and low variance, and a flexible model will
have low bias and high variance. Thus under prediction models, mechanistic
or empirically-driven, this bias-variance trade-off comes into play. For 
mechanistic models, when we add variable to increase model flexibility,
the variance will increase as we reduce the bias. Similarly, empirically-driven
models may seem easier to understand because they are based off of observation
however as a more flexible model, a higher variance will occur in result
of the low-bias nature of the model. 
\
\
**Question 6:**
\
The first question is predictive. We are given a voter profile, and we want 
to predict the outcome of their vote. 
\
The second question is inferential. This question tries to investigate
the relationship between voter and their support on a candidate.There
is no implication of outcome, rather this question looks into what about
a voter's characteristics is important to their candidate support.
\
\
***Exploratory Data Analysis***
```{r}
library(tidyverse)
library(ggplot2)
head(mpg)
```
**Exercise 1:**
```{r}
ggplot(mpg, aes(x=hwy))+
  geom_histogram(color="darkred", fill="red")+
  labs(title="Histogram of Highway MPG", x="Miles per Gallon (Highway)",
       y="Frequency")
```
\
\
This spread is somewhat bi-modal. There are peaks at 17 and 27. This implies
that there are many car models that tend to have 17 mpg and 27 mpg on the 
highway. This bi-modal nature may reflect the different sizes of cars 
like sedans v suvs.\
**Exercise 2:**
```{r}
ggplot(mpg, aes(x=hwy,y=cty))+
  geom_point()+
  labs(title="Scatter Plot of Highway MPG v City MPG", x="MPG on the Highway",
      y= "MPG in the City")
```
There is a clear positive relationship between Highway MPG and City MPG. Cars
that tend to have better highway MPG also have better city MPG. This makes 
sense because a better car, in general, will have better milleage on both
the highway and in the city. 
\
\
**Exercise 3:**
https://stackoverflow.com/a/9231857 (for re-ordering bars)
```{r}
ggplot(mpg, aes(x=reorder(manufacturer, manufacturer, 
                          function(x)-length(x))))+
  geom_bar()+
  coord_flip()+
  labs(title="Barchart of Manufacturers", x="Manufacturers",
         y="Count")
```
Dodge produced the most.\
Lincoln produced the least.
\
\
**Exercise 4:**
```{r}
ggplot(mpg, aes(group= cyl, x=cyl, y=hwy))+
  geom_boxplot()+
  labs(title="Boxplot of Highway MPG grouped by Cylinders", x="Cylinders",
       y="MPG on the Highway")
```
\
We can see from the boxplots, that as the number of cylinders increase, the 
average highway MPG tends to decreases. Which makes sense: Generally more
cylinders means more power but less fuel efficiency.
\
\
**Exercise 5:**
```{r}
library(corrplot)
keeps<-c("displ", "year", "cyl", "cty", "hwy")
mpg_num <-mpg[keeps]
M<-cor(mpg_num)
corrplot(M, type ="lower")
```
The positively correlated pairs are year and displacement, number of cylinder 
and displacement, number of cylinders and year, city mpg
and highway mpg. Some of these make sense. It makes sense that engine 
displacement will improve over time as innovation occurs. The positive
correlation between cylinder count and displacement makes sense because both
contribute the the overall power of the engine. And as mentioned in the 
previous question, it makes sense for highway mpg and city mpg to be positively
correlated because a better car will have better mileage on both the highway
and in the city overall (and vice versa). It surprises me that the number of
cylinders and year is positively correlated. To my limited car knowledge 
cylinder count is vehicle-model specific. That said, in any given year the
number of cylinders var based on the car and have nothing to do with what year
the car is made in (especially considering the time frame of this data).
The negatively correlated pairs are city mpg and displacement, highway mpg
and displacement, city mpg and number of cylinders, highway mpg and
displacement, and city mpg and year. the negative correlations between the 
mileage variables and cylinder variable makes sense. A vehicle with less
cylinders gets better fuel efficiency. Similarly, the negative correlation
between the mileage variables and displacement make sense because lower
displacement uses less power and therefore is more fuel efficient. The negative
correlation between city mpg and year surprises me. I would suspect a positive
correlation between these two variables. It would seem that mpg in the city
would improve overtime with innovation. But this data indicates the opposite.
\
\

